Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-19 PAPER.tex
Updated text to reflect better justification of linear correction between lossy pspec and true pspec

The treatment we outlined in the previous section implicitly assumed that signal loss corrections could be treated in expectation. In practice, however, the signal loss itself is a random variable --- some realizations of the cosmological signal may be more correlated with foreground modes than others, leading to more signal loss. This leads to two issues. The first is that Equation \eqref{eq:LinearPspecSum} may not hold. In particular, a third term that is a quadratic combination of $\f$ and $\s$ (e.g., $\f^\dagger \C^{-1} \s$) may appear. However, we can circumvent this issue by decomposing $\f$ into a sum of two vectors, one that is proportional to $\s$ and one that is orthogonal to $\s$. The former can be absorbed into the EoR term (part of the signal loss estimate), while the latter can be absorbed into the foreground term. 
\dcj{Wait what? I think you are saying essentially that some x can remain in Pout which is what we are seeing. Or is this some extra bit of method which has been added to the analysis...} \acl{I'm not sure that's what I'm saying. (And by ``I'm not sure" I don't mean it in the colloquial sense of ``you're not right and I'm just trying to be polite about it", but rather the literal sense of the phrase ``I'm not sure"). What I was trying to say was that if I wrote Equation \eqref{eq:LinearPspecSum} without the ensemble average, I could say
\begin{equation}
\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.
\end{equation}
(Note that even this is an approximation. In principle there is an infinite series containing all possible powers of $e$ and $f$). Now, we split $f$ into $ f_\parallel + f_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $f_\parallel \equiv \rho e$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. Under this parameterization, one can write the $fe$ term in terms of $e^2$ (which gets absorbed into the other EoR term, plus residual foregrounds that we don't care about).}

The second issue to tackle is how one incorporates the randomness of $\ell_{\rm eor}$ into our signal loss corrections. Phrased in the context of Bayes' rule, we wish to find the posterior probability distribution of $p( \p_{\rm eor} | \phat_{x})$ for $\p_{\rm eor}$ given the uncorrected power spectrum estimate $\phat_{x}$. Since Equation \eqref{eq:Deriv2} is a good approximation of Equation \eqref{eq:Deriv1}, which we argued for in the previous section, we can write $\p_{\rm eor}$ as $P_{in}$, and hence the probability distribution is given by
\begin{equation}
\label{eq:Bayes}
p(P_{in} | \phat_{x}) \propto \mathcal{L} (  \phat_{x} | P_{in})  p(P_{in}),
\end{equation}
where $p(P_{in})$ is the prior on $P_{in}$ and $\mathcal{L} $ is the likelihood function. Since the likelihood is defined as the distribution of the measured result $ \phat_{x}$ given the theoretical power spectrum $P_{in}$, we may construct this function simply by fixing $P_{in}$, and simulating our analysis pipeline for many realizations of the injected EoR signal consistent with this power spectrum. The resulting distribution can be normalized, and the whole process can then be repeated for a different value of $P_{in}$.

@@

The resulting limits for inverse covariance weighted PAPER-64 data (the same data used in Section \ref{sec:Sigloss}, no regularization), using this second signal loss method, are shown in Figure \ref{fig:sigloss_compare} compared with the method used in the main text. Overall the limits show good agreement, with k modes differing by factors of 2 to 4.

>>>>

The treatment we outlined in the previous section implicitly assumed that signal loss corrections could be treated in expectation. In practice, however, the signal loss itself is a random variable --- some realizations of the cosmological signal may be more correlated with foreground modes than others, leading to more signal loss. This leads to two issues. The first is that Equation \eqref{eq:LinearPspecSum} may not hold. In particular, a third term that is a quadratic combination of $\f$ and $\s$ (e.g., $\f^\dagger \C^{-1} \s$) may appear. However, we can circumvent this issue by decomposing $\f$ into a sum of two vectors, one that is proportional to $\s$ and one that is orthogonal to $\s$. The former can be absorbed into the EoR term (part of the signal loss estimate), while the latter can be absorbed into the foreground term. In other words, suppose that we have
\begin{equation}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 + \ell_{\rm eor} |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f} \cdot \mathbf{e}.
\end{equation}
in place of Equation \eqref{eq:LinearPspecSum}. Now, split $\mathbf{f}$ into $\mathbf{ f}_\parallel +\mathbf{ f}_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $\mathbf{ f}_\parallel \equiv \rho \mathbf{e}$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. This then gives
\begin{equation}
\label{eq:fperpfpara}
\phat_x \approx \ell_{\rm fg} |\mathbf{f}|^2 +( \ell_{\rm eor} + \rho  \ell_{\rm fe} ) |\mathbf{e}|^2 + \ell_{\rm fe} \mathbf{f}\cdot \mathbf{f}_\perp.
\end{equation}
Since $|\mathbf{e}|^2$ would be a lossless estimate of the true EoR power spectrum, this expression recovers the linear relation between the true EoR power spectrum and the lossy estimate.
\dcj{Wait what? I think you are saying essentially that some x can remain in Pout which is what we are seeing. Or is this some extra bit of method which has been added to the analysis...} \acl{I'm not sure that's what I'm saying. (And by ``I'm not sure" I don't mean it in the colloquial sense of ``you're not right and I'm just trying to be polite about it", but rather the literal sense of the phrase ``I'm not sure"). What I was trying to say was that if I wrote Equation \eqref{eq:LinearPspecSum} without the ensemble average, I could say
\begin{equation}
\phat_x \approx \ell_{\rm fg} f^2 + \ell_{\rm eor} e^2 + \ell_{\rm fe} fe.
\end{equation}
(Note that even this is an approximation. In principle there is an infinite series containing all possible powers of $e$ and $f$). Now, we split $f$ into $ f_\parallel + f_\perp$, which are the components of the foregrounds that couple to the EoR versus those that do not. Further parameterize $f_\parallel \equiv \rho e$, where $\rho$ is some correlation coefficient that quantifies the degree of correlation between the foregrounds and the EoR. Under this parameterization, one can write the $fe$ term in terms of $e^2$ (which gets absorbed into the other EoR term, plus residual foregrounds that we don't care about).}\acl{Just promoted this last part to something in the main text.}

The second issue to tackle is how one incorporates the randomness of $\ell_{\rm eor}$ into our signal loss corrections. This is manifest in Equation \eqref{eq:fperpfpara}, where the randomness of $\rho$ means that there is an inherent spread in the signal loss correction. Phrased in the context of Bayes' rule, we wish to find the posterior probability distribution of $p( \p_{\rm eor} | \phat_{x})$ for $\p_{\rm eor}$ given the uncorrected power spectrum estimate $\phat_{x}$. Since Equation \eqref{eq:Deriv2} is a good approximation of Equation \eqref{eq:Deriv1}, which we argued for in the previous section, we can write $\p_{\rm eor}$ as $P_{in}$, and hence the probability distribution is given by
\begin{equation}
\label{eq:Bayes}
p(P_{in} | \phat_{x}) \propto \mathcal{L} (  \phat_{x} | P_{in})  p(P_{in}),
\end{equation}
where $p(P_{in})$ is the prior on $P_{in}$ and $\mathcal{L} $ is the likelihood function. Since the likelihood is defined as the distribution of the measured result $ \phat_{x}$ given the theoretical power spectrum $P_{in}$, we may construct this function simply by fixing $P_{in}$, and simulating our analysis pipeline for many realizations of the injected EoR signal consistent with this power spectrum. The resulting distribution can be normalized, and the whole process can then be repeated for a different value of $P_{in}$.

@@

The resulting limits for inverse covariance weighted PAPER-64 data (the same data used in Section \ref{sec:Sigloss}, no regularization), using this second signal loss method, are shown in Figure \ref{fig:sigloss_compare} compared with the method used in the main text. Overall the limits show good agreement, with k modes differing by factors of $\sim$$3$.
